### 2019-05-22 paper/conclusion.tex
conclusion tex

In this work, we explored the back-to-back (B2B) regression, a novel method for
identifying causes from observations building on two successive
multidimensional regressions; from one domain to the other, and then back.
We analyzed the asymptotic behavior of this approach, proving its convergence,
and compared it to several baselines in controlled synthetic experiments. We
showcased our approach on a recent brain imaging dataset, analyzing the timing
of brain responses and their connection to features of words, and obtained
results consistent with prior work in neuroscience literature, confirming the
viability of our method in the case of real data.