### 2019-05-22 paper/conclusion.tex

In this work, we proposed Back-to-Back Regression (B2BR), a method to 
measure the causal influence of a potential set of variables generating some observations.
%
B2BR performs two successive multidimensional regressions: one from the output domain, and another one from the input domain.
%
We provided a theoretical guarantee about the consistency of B2BR, and compared it to several baselines in controlled synthetic experiments.
%
We also applied B2BR to a recent brain imaging dataset, analyzing the timing
of brain responses and their connection to word features.
%
We obtained results consistent with prior work in neuroscience literature, confirming the efficacy of B2BR for real data analysis.