### 2019-05-23 paper/conclusion.tex
conclusion

In this work, we proposed Back-to-Back (B2B) regression, a linear method to
measure the causal influence of a potential set of variables generating multidimensional observations.
%
B2B performs two successive multidimensional regressions: one from the output domain, and another one from the input domain.
%
We provided a theoretical guarantee about the consistency of B2B, and compared it to several baselines in controlled synthetic experiments.
%
We also applied B2B to a recent brain imaging dataset, analyzing the timing
of brain responses and their connection to word features.
%
We obtained results consistent with prior work in neuroscience literature, confirming the efficacy of B2B for real data analysis.
%