### 2020-03-04 paper/__latexindent_temp.tex
clean up

@article{cardinali2013data,
  title={Data assimilation: Observation influence diagnostic of a data assimilation system},
  author={Cardinali, C},
  journal={Shinfield Park, England: European Centre for Medium-Range Weather Forecasts: Research Department},
  year={2013}
}