### 2020-05-19 paper/appendix.tex
E -> S

\newpage
\appendix
\section{Appendix}

\subsection{Proof of consistency theorem}
\label{appendix:theorem_proof}

Proof of the theorem in \ref{sec:theorem}:

\begin{theorem}[B2B consistency - general case]

     Consider the B2B model from equation \ref{eq:model} $$Y = (XS + N)F$$ with
     $N$ centered and full rank noise.

     If $F$ and $X$ are full-rank on $Img(S)$, then, the solution of B2B, $\hat
     H$ minimizes

     $$\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$$ and satisfies

     $$S\hat H = \hat H$$
\end{theorem}
\begin{proof}

 Let $\hat G$ and $\hat H$ be the solutions of the first and second regressions
 of B2B.

 Since $\hat G$ is the least square estimator of $X$ from $Y$
 \begin{align*}
    \hat G = \arg \min_G \mathbb{S}[\left \| YG - X \right \|^2]
\end{align*}
Replacing $Y$ by its model definition $Y = (XS+N)F$, we have
 \begin{align*}
    \hat G &=   \arg \min_G \mathbb{S}[\left \| X - (XS + N)FG \right\|^2] =\arg \min_G \mathbb{S}[\left \| X - XSFG + NFG \right\|^2]
  \end{align*}
  Since $N$ is centered and independent of $X$, we have
  \begin{align}
    	  \hat G &=  \arg \min_G \left \| X - XSFG\right\| ^2  + \left \| NFG\right \| ^2
     \label{eq:Gdoublenorm}
\end{align}

In the same way, for $\hat H$, we have
\begin{align*}
    \hat H = \arg \min_H \mathbb{S}[\| XH - Y \hat{G} \|^2] &=\arg  \min_H \mathbb{S}[\| XH - (XS + N)F \hat G \|^2] \\
    &=\arg \min_H \mathbb{S}[\| X(H - SF \hat G) \| ^2] + \mathbb{S}[\| NF\hat G \| ^2]\\
    &= \arg \min_H \mathbb{S}[\| X(H - SF \hat G) \| ^2]
 \end{align*}
 a positive quantity which reaches a minimum (zero) for
 \begin{align}
    \hat H = SF \hat G
    \label{eq:Hdoublenom}
\end{align}

Let us now prove that $SF\hat G = F\hat G$.

Let $F^\dagger$ be the pseudo inverse of $F$, and $Z=F^\dagger SF\hat G$, we
have $FZ = FF^\dagger SF \hat G$

Since $F$ is full rank on $Img(S)$, we have $FF^\dagger S =S$, and $FZ = SF\hat
G$

As $S$ is a binary diagonal matrix, it is an orthogonal projection and therefore
a contraction, thus
 $$ \| NSF\hat G\|^2 \leq \| NF\hat G \|^2$$ and
 $$\left \| X - XSFZ\right \| ^2  + \left \| NFZ\right \| ^2 = \| X - XSF\hat G \| ^2  + \| NSF\hat G \| ^2 \leq \| X - XSF\hat G \| ^2  + \| NF\hat G \| ^2$$

But since $\hat G =  \arg \min_G \left \| X - XSFG\right\| ^2  + \left \| NFG\right \| ^2$, we also have
$$\left \| X - XSF\hat G\right\| ^2  + \left \| NF\hat G\right \| ^2 \leq \left \| X - XSFZ\right \| ^2  + \left \| NFZ\right \| ^2$$

Summarizing the above,
$$\left \| X - XSF\hat G\right\| ^2  + \left \| NF\hat G\right \| ^2 \leq \| X - XSF\hat G \| ^2  + \| NSF\hat G \| ^2 \leq \| X - XSF\hat G \| ^2  + \| NF\hat G \| ^2$$
$$\left \| X - XSF\hat G\right\| ^2  + \left \| NF\hat G\right \| ^2 = \| X - XSF\hat G \| ^2  + \| NSF\hat G \| ^2$$
$$\left \| NF\hat G\right \| ^2 =  \| NSF\hat G \| ^2$$

$N$ being full rank, this yields $SF\hat G = F\Hat G$.

Replacing into $\eqref{eq:Gdoublenorm}$, and setting $H = SFG$, we have
\begin{align*}
	\hat G &=  \arg \min_G  \left \| X - XSFG\right \| ^2  + \left \| NFG\right \| ^2 \\
	&=   \arg \min_G \left \| X - XSFG\right \| ^2  + \left \| NSFG\right \| ^2 \\
	\hat H &=  \arg \min_H \left \| X - XH\right \| ^2  + \left \| NH\right \| ^2
	\label{eq:4}
\end{align*}

Finally, $S\hat H = S SF\hat G = SF\hat G = \hat H$, since $S$, a binary
diagonal matrix, is involutive. This completes the proof.
\end{proof}

\newpage
\subsection{Modeling measurement noise}

Equation \ref{eq:model} does not explicitly contain a measurement noise term.
Yet, in most experimental cases, the problem is best described as:
\begin{equation}
  Y = (XS+N)F+M
  \label{eq:model_noisy_measure}
\end{equation}
  with $M\in \mathbb{R}^{n \times d_y}$.

This equation is actually equivalent to Equation \ref{eq:model} given our
hypotheses. Indeed, we can rewrite $M = MF^{-1}F$ over $Img(F)$, which leads to:
$$Y = (XS+N)F+M = (XS+N+MF^{-1})F = (XS+N')F$$

Consequently, assuming that $F$ is full rank on $Img(XS)$, B2B yields the same
solutions to equations \ref{eq:model} and \ref{eq:model_noisy_measure}.

\subsection{Feature importance}
\label{appendix:feature_importance}

For B2B, feature importance is assessed as follows:

\begin{algorithm}[H]
    %\SetAlgoLined
    \KwIn{$X_{train} \in \mathbb{R}^{n \times d_x}$, $X_{test} \in
\mathbb{R}^{n' \times d_x}$, $Y_{train} \in \mathbb{R}^{n\times d_y}$, $Y_{test}
\in \mathbb{R}^{n'\times d_y}$, } \KwOut{estimate of prediction improvement
$\Delta{R} \in \mathbb{D}^{d_x}$.} $H, G = \text{B2B}(X_{train}, Y_{train})$\;
$R_{full} = \text{corr}(X_{test} H, Y_{test} G)$\;

    \For{$i = 1, \ldots, d_x$}{ $K = Id$\; $K[i] \leftarrow 0$\;
    $R_{k} =
\text{corr}(X_{test} K H, Y_{test} G_i)$\; $\Delta R_i = R_{full} - R_{k}$\; }
\Return{$\Delta R$} \caption{B2B feature importance.} \label{algorithm:b2b_fi}
\end{algorithm}

For the Forward Model, the feature importance is assessed as follows:

\begin{algorithm}[H]
    %\SetAlgoLined
    \KwIn{$X_{train} \in \mathbb{R}^{n \times d_x}$, $X_{test} \in
\mathbb{R}^{n' \times d_x}$, $Y_{train} \in \mathbb{R}^{n\times d_y}$, $Y_{test}
\in \mathbb{R}^{n'\times d_y}$, } \KwOut{estimate of prediction improvement
$\Delta{R} \in \mathbb{D}^{d_x, d_y}$.}

$H = \text{LinearRegression}(X_{train}, Y_{train})\;
R_{full} = \text{corr}(X_{test} K, Y_{test})$\;

    \For{$i = 1, \ldots, d_x$}{ $K = Id$\; $K[i] \leftarrow 0$\; $R_{k} = \text{corr}(X_{test}
K H, Y_{test})$\; $\Delta R_i = R_{full} - R_{k}$\; } \Return{$\Delta R$}
\caption{Forward feature importance.} \label{algorithm:fwd_fi} \end{algorithm}

% \iffalse
% \begin{enumerate}
%   \item Fit $H$ given $X_{train}$ and $Y_{train}$
%   \item For each feature $i$:
%   \begin{itemize}
%     \item Define $K$, and identity matrix whose row $i$ has been zeroed-out
%     \item Fit $H^i_k$ given $X_{train} K$ and $Y_{train}$
%   \end{itemize}
%   \end{enumerate}
% \fi

For the CCA and PLS models, the feature importance is assessed as follows:

\begin{algorithm}[H]
    %\SetAlgoLined
    \KwIn{$X_{train} \in \mathbb{R}^{n \times d_x}$, $X_{test} \in
\mathbb{R}^{n' \times d_x}$, $Y_{train} \in \mathbb{R}^{n\times d_y}$, $Y_{test}
\in \mathbb{R}^{n'\times d_y}$, } \KwOut{estimate of prediction improvement
$\Delta{R} \in \mathbb{D}^{d_x, d_z}$.} $H, G = \text{CCA}(X_{train},
Y_{train})$\; $R_{full} = \text{corr}(X_{test} H, Y_{test} G)$\;

    \For{$i = 1, \ldots, d_x$}{ $K = Id$\; $K[i] \leftarrow 0$\; $R_{k} = \text{corr}(X_{test} K H,
Y_{test} G)$\; $\Delta R_i = R_{full} - R_{k}$\; } \Return{$\Delta R$}
\caption{CCA and PLS feature importance.} \label{algorithm:cdp_fi}
\end{algorithm}

For the Backward Model, feature importance cannot be assessed because there is
no prediction.

\subsection{Recovering S}
\label{recovering}

In case of noise, B2B yields non binary $\hat S$. Three thresholding rules can
be used to binarize its values thus explicitly recover "causal" features.

First, given known signal-to-noise ratio, the threshold above which a feature
should considered to be "causal" can be derived analytically. Indeed, Equation
$\ref{eq:diagk}$ implies that the $k$ first diagonal elements of $\hat H$ are
bounded:  % we assume that $F$ has full rank over $Img(S)$.

$$0 \leq \frac{\sigma_{X_k}}{\sigma_{X_k} +\sigma_{N_1}} \leq
\text{diag}_k(\hat{H})\leq  \frac{\sigma_{X_1}}{\sigma_{X_1} +\sigma_{N_k}}$$
where $\sigma_{X_1}$, $\sigma_{X_k}$, $\sigma_{N_1}$ and $\sigma_{N_k}$ denote
the largest and smallest eigenvalues of $\Sigma_{X_1 X_1}$ and $\Sigma_{N_1
N_1}$.

The average value $\mu$ of non-zero coefficients of $\text{diag}(\hat H)$ is the
trace of $\hat H$ divided by $k$, and can be computed as
\begin{equation}
\mu = \frac{Var(X)}{Var(X)+Var(N)}
\end{equation}

The decision threshold between "causal" and "non-causal" elements is thus a fraction
$\mu$, whose proportion arbitrarily depends on the necessity to favor type I and
type II errors. In practice, we cannot use this procedure for our MEG study,
because signal-to-noise ratio is unknown.

Second, $\text{diag}(\hat H)$ can be binarized with the Sonquist-Morgan
criterion \citep{sonquist_morgan}, a non-parametric clustering procedure
separating small and large values in a given set. This procedure maximizes the
ratio of inter-group variance while minimizing the intra-group variance, over
all possible splits of the diagonal into $p$ largest values and $d_x-p$ smallest
values. Let $m_0$ and $m_1$ be the average values of the two clusters, $p$ and
$d_x-p$ their size, and $v$ the total variance of the sample, Sonquist-Morgan
criterion maximizes \citep{kass75}:
\begin{equation}
  \frac{p(d_x-p)}{d_x} \frac{(m_1 - m_0)^2}{v}
\end{equation}

This procedure assumes that there exists at least one causal and at least one
non-causal feature.
%
Third, second-order statistics across multiple datasets can be used to identify
the elements of $\text{diag}(\hat H)$ that are significantly different from 0.
This procedure is detailed in the method section of our MEG experiment.

Overall, these three procedures thus vary in their additional assumptions: i.e.
(1) a known signal-to-noise ratio, (2) the existence of both causal and
non-causal factors or (3) independent repetitions of the experiment.

\section{Additional Figures}

% \begin{figure}[t!]
%   \centering
%   \includegraphics[width=\textwidth, trim=0cm 0cm 0cm 0cm, clip=True]{figures/meg_supp.pdf}
%   \caption{Average $\Delta R$ for each $Y$ dimension (Forward model), canonical components (PLS and CCA) or feature (B2B) across subjects. Formally, Backward model cannot have a $\Delta R$ because it never unmixes the multiple $X$ features. In such cases, we thus simply report the decoding score.}
%   \label{fig:meg_supp}
% \end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/delta_auc.pdf}
  \caption{Synthetic experiments. Distribution (over conditions) of AUC (top) and Feature
  Importance $\Delta R$ (bottom) metrics between our method (y-axis) and the baselines
  (x-axis). Each dot is a distinct synthetic experiment. Dots below the diagonal indicates
  that B2B outperform the tested model. \label{fig:auc_plots}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{figures/duration}
  \caption{Wall-clock run-time for our method B2B and for the baselines. Each dot is a
  distinct synthetic experiment. B2B runs much faster than cross-decomposition baselines.
  \label{fig:duration}}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\textwidth,
                     trim=0cm 0cm 0cm 0cm,
                     clip=True]{figures/fmri_decod.pdf}
    \includegraphics[width=0.7\textwidth,
                     trim=0cm 0cm 0cm 0cm,
                     clip=True]{figures/fmri_betas.pdf}

    \label{fig:fmri_supp}
  \end{center}
  \caption{Top. P-values across subjects derived from a Backward decoding of
  each feature (columns). Bottom. P-values across subjects derived from
  the $\beta$ coefficients of a Forward Model and the $\hat S$ of an unbiased
  B2B (i.e. B2B where $H$ is not regularized). Each cell displays the left
  hemisphere view laterally (top) or ventrally (bottom)}
\end{figure}

\subsection{Robustness to increasing number of factors}

To test whether each of the methods robustly scales to an increasingly
large number of potential causes $X$, we enhanced the four ad-hoc features
(word length, word frequency, word function, dummy variable) with another
ten features. These additional features corresponds to the first dimensions
of word embedding as provided by Spacy \citep{spacy2}. The results shown
in Figure \ref{fig:embeddings}, show that the feature importance of ad-hoc
features as derived by B2B remain unchanged and are actually improved.

\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/compare_embeddings.pdf}
  \caption{Comparison of $\Delta R$ when the models are tested on four
  variables (top) and when the models are tested on an these four variables
  as well as another 10 word-embedding features (bottom). These results
  illustrate that, unlike Regularized CCA, B2B remains robust even when
  the number of tested factors increases.
  \label{fig:embeddings}}
\end{figure}