### 2020-05-20 paper/paper.tex

\documentclass[preprint,12pt,3p]{elsarticle}

\usepackage{lineno,hyperref}
% \modulolinenumbers[5]

\usepackage{amsmath, amsfonts, microtype, xcolor, tikz, graphicx, amsthm}
\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage{times,wrapfig}
\usepackage{cleveref}

\DeclareMathOperator*{\argmax}{argmax} % thin space, limits underneath in displays

\newtheorem{theorem}{Theorem}
\newtheorem*{theorem*}{Theorem}

\SetKwComment{Comment}{$\triangleright$\ }{}

\usetikzlibrary{calc}

\tikzset{
    ncbar angle/.initial=90,
    ncbar/.style={
        to path=(\tikztostart)
        -- ($(\tikztostart)!#1!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztotarget)$)
        -- ($(\tikztotarget)!($(\tikztostart)!#1!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztotarget)$)!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztostart)$)
        -- (\tikztotarget)
    },
    ncbar/.default=0.5cm,
}

\tikzset{round left paren/.style={ncbar=0.5cm,out=110,in=-110}}
\tikzset{round right paren/.style={ncbar=0.5cm,out=70,in=-70}}

% \journal{Neuroimage}

% \bibliographystyle{elsarticle-num}
\bibliographystyle{plain}

\begin{document}
\begin{frontmatter}

\title{Discriminating the Influence of Correlated Factors \\from Multivariate Observations: the Back-to-Back Regression}

\author[1,2]{Jean-Rémi King\corref{email}}
\cortext[email]{corresponding author: jeanremi@fb.com}

\author[2]{Fran\c{c}ois Charton}
\author[2]{David Lopez-Paz}
\author[2]{Maxime Oquab}
\address[1]{Laboratoire des systèmes perceptifs, PSL University, CNRS}
\address[2]{Facebook AI}

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate independently and ii)
observations are multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a linear method designed to efficiently
estimate, from a set of correlated factors, those that most plausibly account for
multidimensional observations. First, we prove the consistency of B2B, its links to
other linear approaches, and show how it can provide a robust, unbiased and
interpretable scalar estimate for each factor.
Second, we use a variety of simulated data to show that B2B outperforms
forward modeling ("encoding"), backward modeling ("decoding") as well as
cross-decomposition modeling (i.e.. canonical
correlation analysis and partial least squares) on causal identification when
the factors and the observations are not orthogonal.
Finally, we apply B2B to a hundred magneto-encephalography recordings and to a hundred functional Magnetic
Resonance Imaging recordings acquired during a one hour reading task. B2B successfully disentangles
the respective contribution of collinear factors such as word length, word
frequency in the early visual and late associative cortical
responses respectively. These factors are better disentangled with B2B than with
other standard techniques. We discuss how the speed and generality of B2B
sets the foundations of an omnibus test for disentanglement of covarying
factors in multidimensional observations.
\end{abstract}

\begin{keyword}
  Feature Discovery \sep
  MEG \sep
  fMRI \sep
  Decoding \sep
  Encoding \sep
  Cross-Decomposition \sep
  Reading \sep
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}
\input{introduction.tex}

\section{Back-to-Back regression}
\label{sec:algorithm}

\begin{figure}[t!]

  \centering
  \begin{tikzpicture}
  \newcommand\posY{0}
  \newcommand\posX{3}
  \newcommand\posS{5}
  \newcommand\posN{7}
  \newcommand\posF{10}

  \node[thick, draw=black, minimum height=3cm, minimum width=2cm, fill=yellow!10] (Y) at (\posY, 0){};
  \node[] (eq) at (1.5, 0){$=$};
  \node[] (times) at (4, 0){$\times$};
  \node[thick, draw=black, minimum height=3cm, minimum width=1cm, fill=yellow!10] (X) at (\posX, 0){};
  \node[thick, draw=black, minimum height=1cm, minimum width=1cm, fill=red!10] (S) at (\posS, 0){};

  \draw[fill=white] (\posS - 0.5 + 0.0, 0.5 - 0.0) rectangle (\posS - 0.5 + 0.0 + 0.1, 0.5 - 0.0 - 0.1);
  \draw[fill=black] (\posS - 0.5 + 0.1, 0.5 - 0.1) rectangle (\posS - 0.5 + 0.1 + 0.1, 0.5 - 0.1 - 0.1);
  \draw[fill=white] (\posS - 0.5 + 0.2, 0.5 - 0.2) rectangle (\posS - 0.5 + 0.2 + 0.1, 0.5 - 0.2 - 0.1);
  \draw[fill=white] (\posS - 0.5 + 0.3, 0.5 - 0.3) rectangle (\posS - 0.5 + 0.3 + 0.1, 0.5 - 0.3 - 0.1);
  \draw[fill=white] (\posS - 0.5 + 0.4, 0.5 - 0.4) rectangle (\posS - 0.5 + 0.4 + 0.1, 0.5 - 0.4 - 0.1);
  \draw[fill=white] (\posS - 0.5 + 0.5, 0.5 - 0.5) rectangle (\posS - 0.5 + 0.5 + 0.1, 0.5 - 0.5 - 0.1);
  \draw[fill=white] (\posS - 0.5 + 0.6, 0.5 - 0.6) rectangle (\posS - 0.5 + 0.6 + 0.1, 0.5 - 0.6 - 0.1);
  \draw[fill=black] (\posS - 0.5 + 0.7, 0.5 - 0.7) rectangle (\posS - 0.5 + 0.7 + 0.1, 0.5 - 0.7 - 0.1);
  \draw[fill=white] (\posS - 0.5 + 0.8, 0.5 - 0.8) rectangle (\posS - 0.5 + 0.8 + 0.1, 0.5 - 0.8 - 0.1);
  \draw[fill=black] (\posS - 0.5 + 0.9, 0.5 - 0.9) rectangle (\posS - 0.5 + 0.9 + 0.1, 0.5 - 0.9 - 0.1);

  \node[] (plus) at (6, 0){$+$};
  \node[thick, draw=black, minimum height=3cm, minimum width=1cm, fill=blue!10] (N) at (\posN, 0){};
  \node[thick, draw=black, minimum height=1cm, minimum width=2cm, fill=red!10] (F) at (\posF, 0){};
  \draw [thick] (2.25, -1.5) to [round left paren ] (2.25, 1.5);
  \draw [thick] (7.75, -1.5) to [round right paren ] (7.75, 1.5);

  \node[] (times2) at (8.5, 0){$\times$};

  \node[] (annY) at (\posY, -1.8){\scalebox{0.85}{$Y \in \mathbb{R}^{m \times d_y}$}};
  \node[] (annX) at (\posX, -1.8){\scalebox{0.85}{$X \in \mathbb{R}^{m \times d_x}$}};
  \node[] (annS) at (\posS, -1.8){\scalebox{0.85}{$S \in \mathbb{D}^{d_x \times d_x}$}};
  \node[] (annN) at (\posN, -1.8){\scalebox{0.85}{$N \in \mathbb{R}^{m \times d_x}$}};
  \node[] (annF) at (\posF, -1.8){\scalebox{0.85}{$F \in \mathbb{R}^{d_x \times d_y}$}};

  \node[] (labY) at (\posY, 2){Observations};
  \node[] (labX) at (\posX, 2){Factors};
  \node[] (labS) at (\posS, 2){Cause};
  \node[] (labS) at (\posS, 1.7){selection};
  \node[] (labN) at (\posN, 2){Noise};
  \node[] (labF) at (\posF, 2){Cause-effect};
  \node[] (labF) at (\posF, 1.7){mapping};

  \node[] (sim1) at (0,-3.25) {\scalebox{0.85}{$X \sim P(X)$}};
  \node[] (sim2) at (0,-3.65) {\scalebox{0.85}{$N \sim P(N)$}};
  \node[] (reg1) at (6,-3.25) {$\hat{S} = \text{diag}(\underbrace{(X_2^\top X_2 + \Lambda_X)^{-1} X_2^\top Y_2\overbrace{(Y_1^\top Y_1 + \Lambda_Y)^{-1} Y_1^\top X_1}^{\text{1) } \hat{X} : \text{ regression from } Y \text{ to } X}}_{\text{2) regression from } X \text{ to } \hat{X}})$};
  \end{tikzpicture}
  \caption{Back-to-back regression identifies the subset of factors $S_{ii} = 1$ in $X$ that influence $m$ multidimensional observations $Y$ by 1) regressing from $Y$ to $X$ to obtain $\hat{X}$, and 2) returning the diagonal of the regression coefficients from $X$ to $\hat{X}$.}
\label{fig:b2b}
\end{figure}

\subsection{Problem setup}

We consider the measurement of multivariate signal $Y \in \mathbb{R}^{m \times
d_y}$ (the dependent variables, e.g. the neural responses),
generated from a set of putative causes $X \in \mathbb{R}^{m \times
d_x}$ (the independent variables, e.g. the features of a stimulus),
via some unknown linear apparatus $F \in \mathbb{R}^{d_x \times d_y}$ (e.g. the
projection of neural activity onto MEG channels or fMRI voxels).
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $S
\in \mathbb{D}^{d_x \times d_x}$, we denote by $XS$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    y_i = (x_i S + n_i) F
    \label{eq:model}
\end{equation}
%
where $i$ is a given sample, and $n_i$ is a sample-specific additive noise drawn from a
centered distribution.
%
While $X$ and $N$ are independent, we allow each of
them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $m$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is
illustrated in Fig.~\ref{fig:b2b}.
%
Given the model in Equation~\cref{eq:model}, the goal of Back-to-Back
Regression (B2B) is to estimate the matrix $S$, i.e. to identify the factors
that reliably account for the multivariate observations.

\subsection{Algorithm}

Back-to-Back Regression (B2B) consists of two steps.
%
First, we estimate the linear regression coefficients $\hat G$ from $Y$ to $X$,
and construct the predictions $\hat X = Y \hat G$.
%
This backward regression recovers the correlations between $Y$ and each factor
of $X$.
%
Second, we estimate the linear regression coefficients $\hat H$ from $X$ to
$\hat X$.
%
The diagonal of the regression coefficients $\hat H$, denoted by $\hat{S} =
\text{diag}(\hat{H})$, is the desired estimate of the causal influence matrix
$S$, as detailed in the \ref{appendix:theorem_proof}.

If using l2-regularized least-squares \citep{hoerl1959optimum, rifkin2007notes},
B2B has a closed form solution:
\begin{align}
    \hat G &= (Y^\top Y + \Lambda_Y)^{-1} Y^\top X,\label{eq:solG}\\
    \hat H &=(X^\top X + \Lambda_X)^{-1} X^\top Y \hat G,\label{eq:solH}
\end{align}
%
where $\Lambda_X$ and $\Lambda_Y$ are two diagonal matrices of regularization
parameters, useful to invert the covariance matrices of $X$ and $Y$ if these are
ill-conditioned.

Performing two regressions over the same data sample can result in overfitting,
as spurious correlations in the data absorbed by the first regression will be
leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ into two disjoint sets $(X_1, Y_1)$
and $(X_2, Y_2)$.
%
The first regression is performed using $(X_1, Y_1)$, and the second
regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, the two
successive regressions are
repeated over many random splits, and the final estimate $\hat S$ of the causal
influence matrix is the average over the estimates associated with each split
\citep{breiman1996bagging}.
%
To accelerate this ensembling procedure, we use an efficient
leave-one-out (LOO) cross-validation scheme as detailed in \citep{rifkin2007notes}
% http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf,
as follows:
%

\begin{equation}
\hat{Y}_{LOO} = (\Sigma_X G Y - \text{diag}(\Sigma_X G) Y) \;/\; \text{diag}(I - \Sigma_X G) \qquad \text{(element-wise division)}
\end{equation}

where $\Sigma_X$ is the $X$ kernel matrix and where $G$ is computed with an
eigen decomposition of $X$:
\begin{equation}
  \begin{aligned}
  \Sigma_X = Q V Q^T \\
  G = Q (V + \lambda I)^{-1} Q^T\\
  \end{aligned}
\end{equation}
where $Q$, $V$ and $\lambda$ are the eigen vectors, eigen values and
regularization, respectively.

We summarize the B2B procedure in Algorithm~\ref{algorithm:b2br}.
%
The rest of this section provides a theoretical guarantee on the correctness of
B2B.

% JR: shall we update algorithm with LOO trick?
\begin{algorithm}[H]
    %\SetAlgoLined
    \KwIn{input data $X \in \mathbb{R}^{m \times d_x}$, output data $Y \in \mathbb{R}^{m\times d_y}$, number of repetitions $m \in \mathbb{N}$.}
    \KwOut{estimate of causal influences $\hat{S} \in \mathbb{D}^{d_x \times d_x}$.}
    $\hat{S} \leftarrow 0$\;
    \For{$i = 1, \ldots, m$}{
        $(X, Y) \leftarrow \text{ShuffleRows}((X, Y))$\;
        $(X_1, Y_1), (X_2, Y_2) \leftarrow \text{SplitRowsInHalf}((X, Y))$\;
        $\hat{G} = \text{LinearRegression}(Y_1, X_1)$ \Comment*[r]{$\hat G = (Y_1^\top Y_1 + \Lambda_Y)^{-1} Y_1^\top X_1$}
        $\hat{H} = \text{LinearRegression}(X_2, Y_2 \hat{G})$ \Comment*[r]{$\hat H=(X_2^\top X_2 + \Lambda_X)^{-1} X_2^\top Y_2 \hat G$}
        $\hat{S} \leftarrow \hat{S} + \text{diag}(\hat{H})$\;
    }
    $\hat{S} \leftarrow \hat{S} / m$\;
    $\hat{W} \leftarrow \text{LinearRegression}(X \hat{S}, Y)$\;
    \Return{$\hat{S}$, $\hat{W}$}
    \caption{Back-to-back regression.}
    \label{algorithm:b2br}
\end{algorithm}

\subsection{Theoretical guarantees}
\label{sec:theorem}

\begin{theorem}[B2B consistency - general case]

     Consider the B2B model from Equation $Y = (XS + N)F$, $N$ centered and full
     rank noise.
     %
     Let $Img(M)$ refers to the image of the matrix $M$. If $F$ and $X$ are
     full-rank on the $Img(S)$, then, the solution of B2B, $\hat H$, will
     minimize
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $S\hat H = \hat H$
\end{theorem}
%
\begin{proof}
  See Appendix \ref{appendix:theorem_proof}.
\end{proof}

Since  $S\hat H = \hat H$, we have
\begin{equation}
  \hat H = \arg \min_H  \left \| X - XSH\right\| ^2  + \left \| NSH\right \| ^2 = (S X^\top XS +SN^\top NS) ^\dagger SXX^\top.
\end{equation}
% So, $\hat H = \arg \min_H  \left \| X - XSH\right\| ^2  + \left \| NSH\right \| ^2 = (S X^\top XS +SN^\top NS) ^\dagger SXX^\top$.

Assuming, without loss of generality, that the active features in $S$ are the $k
\in \mathbb{Z}: k \in [0, d_x]$ first features, and rewriting $X=(X_1,X_2)$ and
$N=(N_1,N_2)$ ($X_1$ and $N_1$ containing the $k$ first features), we have:

% $$
\begin{equation}
  X^\top X = \left(\begin{array}{lccl}\Sigma_{X_1 X_1} & \Sigma_{X_1 X_2} \\ \Sigma_{X_1 X_2} & \Sigma_{X_2 X_2}\end{array}\right),\qquad\qquad N^\top N = \left(\begin{array}{lccl}\Sigma_{N_1 N_1} & \Sigma_{N_1 N_2} \\ \Sigma_{N_1 N_2} & \Sigma_{N_2 N_2}\end{array}\right),
\end{equation}

where $\Sigma_{A B}$ is the covariance of $A$ and $B$, and:
% and
\begin{equation}
  \hat H = \left(\begin{array}{cc}(\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1} & (\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_2} \\0 & 0\end{array}\right)
\end{equation}
\begin{equation}
  \text{diag}_k (\hat H) = \text{diag}((\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1}) = \text{diag}((I+\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1})^{-1})
  \label{eq:diagk}
\end{equation}
%
% $$
% $$
% $$
% $$
%
In the absence of noise, we have $\Sigma_{N_1 N_1}=0$, and so
$\text{diag}_k(\hat H)=I$, and $$\text{diag}(\hat H) = \text{diag}(S)$$
Therefore, we recover $S$ from $\hat H$.

In the presence of additive noise, the causal factors of $S$ correspond to the positive
elements of $\text{diag}(\hat H)$. The methods to recover them are presented in
the Appendix (\ref{recovering}).

Note that $hat \S $ is unbiased, in the sense that it is centered around zero when there is no
effect, only if the second regression $H$ is not regularized. Second-level
statistics testing whether $hat \S$ is superior to 0 are thus only valid if $H$
is not regularized.

\section{Experiments}

We perform three sets of experiments to evaluate B2B: one on controlled synthetic
data, a second one on a real, large-scale functional Magnetic Resonance Imaging
(fMRI) dataset and a third one on a real, large-scale magneto-encephalography (MEG)
dataset.
%
We use scikit-learn's PLS and RidgeCV \citep{sklearn} as well as Pyrcca's
regularized canonical component analysis (RegCCA, \citep{bilenko2016Pyrcca})
objects to compare B2B against the standard baselines, with common
hyper-parameter optimizations.

\subsection{Synthetic Experiment}
\label{sec:experiment_synthetic}
\input{synthetic.tex}

\subsection{functional Magnetic Resonance Imaging Experiment}
\label{sec:experiment_fmri}
\input{fmri_exps.tex}

\subsection{Magneto-encephalograph Experiment}
\label{sec:experiment_real}
\input{meg_exps.tex}

\section{Discussion}

Here, we introduce B2B, a method to disentangle the causal contribution of
collinear factors from multidimensional observations. After proving the validity
of B2B, we show that it generally compares favorably against baseline models
both on a wide spectrum of synthetic data and on two large neuroimaging datasets.

In addition, B2B can be very fast to compute as long as both $H$ and $G$ are based
on l2 optimization with l2 regularization (Supplementary
Fig.~\ref{fig:duration}), as is done in the above experiments. However, B2B
does not need not be limited to l2 optimization and regularization: the $H$ and $G$ operators
can in principle be found with e.g. support vector or random forest regression.
We reserve the study of these computational variants for future work.

Like forward and cross-decomposition models, B2B is limited by the correlations
between factors. At the extreme, if two factors are identical, and thus fully
correlated, no statistical method can disentangle their relative causal
contribution, and intervention is thus mandatory. In practice, this implies that, like
general linear models, B2B will best work with high signal-to-noise ratio and/or
orthogonal factors, and will see its sensitivity diminish when the
signal-to-noise ratio of collinear factors drops.

In the present neuroimaging context, B2B follows a long series of statistical
methods designed to characterize brain representations - i.e. to identify what
sensory feature causes specific brain responses \citep{naselaris2011encoding}.
In this regard, CCA and PLS have been used in electrophysiology and neuroimaging
to track representations (e.g. \citep{musch2020transformation}) as well as
to denoise recordings as well as to align subjects \citep{cca_hotelling, de2019multiway}. While
CCA and PLS relates to B2B, these methods diverge in several important ways.
First, they have different objectives: CCA aims to find the potentially numerous and poorly
interpretable components where $X$ and $Y$ are maximally correlated, whereas B2B
aims to recover the causal factors from X to Y. Second, B2B is not symmetric
between $X$ and $Y$: it aims to identify specific causal features by first
optimizing over the decoders $G$ and then over $H$. By contrast, CCA and PLS are
symmetric between $X$ and $Y$, and aims to find $G$ and $H$ jointly such that they
project $X$ and $Y$ on maximally correlated dimensions. Third, CCA is based an
eigen decomposition of $XH$ and $YG$ - the corresponding canonical components
are thus mixing the $X$ features in way that limit interpretability and
potentially dilute the impact of each feature onto multiple components. In
contrast B2B assesses each feature $X_i$ on a single $Y$ component specifically
selected to maximize signal-to-noise ratio of that feature $i$. Fourth CCA does
not separately optimize two distinct regularization parameters for $G$ and $H$,
whereas B2B does. Finally, CCA does not use different data splits to
estimate $G$ and $H$. Together, these differences may explain why B2B generally
outperform CCA on estimating causal influences (Figs.~\ref{fig:percondition} and
\ref{fig:auc_plots}).

One popular method to investigate multidimensional patterns of brain activity is
Representational Similarity Analysis (RSA)
\citep{kriegeskorte2008representational}. RSA quantifies the
similarity of brain responses associated with specific categorical conditions
(e.g. distinct images), by (1) fitting one-against-all classifiers on each
category and (2) testing whether these classifiers discriminates all other
categories. The resulting categories $\times$ categories confusion matrix is
then analyzed, generally in an unsupervised manner, to reveal the categories
that present similar brain activity patterns. B2B
subsumes RSA in that (1) it can use regressions instead of one-hot
classifications and (2) it is fully supervised.
Consequently, and unlike RSA, B2B (1) provides interpretable coefficients and
(2) can generalize to new items and new contexts. In practice, these elements
allow B2B to apply to event-related paradigms and latent variable analyses,
whereas RSA can only be applied when the same one-hot-encoded condition is
repeated multiple times.

More generally, the present empirical results, together with their theoretical
foundations, suggest that B2B may serve as a method of choice for causal
discovery when feature-wise intervention is limited.

\section{Acknowledgements}
We are thankful to Gael Varoquaux and Alexandre Gramfort for the valuable
feedback. This work was supported by ANR-17-EURE-0017 and the Fyssen Foundation
to JRK for his work at PSL.

\bibliography{paper}

\newpage
\clearpage
\section{Appendices}
\input{appendix}

\end{document}