### 2020-05-20 paper/response.tex
clean up

Reviewers' comments:

Reviewer #1:
    There is one major limitation for the broader application of this approach and
    that's demonstrating its effectiveness in fMRI data. The majority of
    neuroimaging studies are done using fMRI (and I do not say this to discount the
    other valuable tools of EEG, MEG, NIRS, etc.). To get other neuroscientists to
    grab onto this approach, I recommend the authors perform another analysis using
    fMRI data. My suggestion for the authors would be to use data from the Human
    Connectome Project (http://www.humanconnectomeproject.org/). The outcome
    variable used with this approach is not important as so much the demonstration
    that their technique is effective for fMRI data. Without including fMRI data,
    the approach comes off as useful but speculative for fMRI data, which
    drastically limits its impact potential.

We thank Reviewer #1 for this remark. We now apply the same analyses to
fMRI recordings of 100 subjects performing a reading experiment similar to the
MEG study described in original manuscript. We kept the exact same methods and parameters, and
used the fmriprep package to perform the preprocessing.

Overall, the results consistently show that B2B compares favorably to other methods.
Furthermore, these novel analyses complements the temporal results obtained with
MEG. Specifically, the effects of Word Length are predominantly observed over the early visual
cortices, and the Word Frequency are predominantly observed over the fronto-temporal areas.

# Reviewer 2:
    First, the authors justify their approach of using backward modeling by the
    ability to estimate common noise sources. However, this would also be possible
    using multivariate forward modeling. For that reason, I'm wondering how
    multivariate forward modeling would fare under these circumstances. I would find
    it highly informative to include a (similarly cross-validated) multivariate GLM
    to the analyses and results.

We thank Reviewer 2 for this remark. The multivariate GLM / MANOVA is indeed
a relevant analysis. We now indicate in supplementary materials the following analyses:

"
Following the recommendations of one of our reviewers, we implemented a
multivariate variant of the forward model, i.e. a MANOVA, using the statsmodels
implementation \citep{seabold2010statsmodels}. MANOVA is primarily used as a
inferential statistics, and does not trivially convert to a predicting method.
Consequently, we did not find a way to compare MANOVA against B2B with the
$\Delta R$ evaluation. However, the effects of MANOVA are generally summarized
with the Wilk's Lambda stastistics or its transformation into an $F$-value.
For each searchlight, we thus use the $F$-values of the
Wilk's Lambda statistics as proxy for $\hat S$ and feeds it to a second-level
Wilcoxon signed-rank tests across subjects, like we did for the other models.

The results show that all factors, including the Dummy variable, are
systematically above chance level in all recorded brain regions (Fig.~\ref{fig:fmri_supp}). This result
can be explained by the large dimensionality of $Y$. Indeed, limiting the
searchlight to a 1mm radius did not lead to these spurious effects but provided
results similar to the Forward model.

Nonetheless, MANOVA does appear to capture some plausible effects. Indeed, the
$F$-values obtained for both Word Length and Word Frequency were weakly but
significantly higher than those obtained with the Dummy variable in the
occipital and temporal brain areas (Fig.~\ref{fig:fmri_supp}). This results
suggests that the effect size of the MANOVA can be biased and, thus, is not
valid for second-level statistics.

Overall, this suggests that MANOVA (1) can lead to positively biased estimates of
$\hat S$ (2) appears weaker than B2B in terms of second-level analysis across
subjects (3) misses the effect of Word Function detected with the Forward model,
and (4) does not trivially translate into a prediction tool. Together these
elements thus suggest that MANOVA is less suitable to the present objective than
B2B.
"

> Likewise, while I applaud the authors efforts in conducting these analyses,
    I am wondering whether the impact of this method could not be made even
    stronger by adding more experimental findings. Are there any other openly
    available datasets that could be used to demonstrate the ability to
    disentangle known factors?

We agree with Reviewer 2. As indicated to Reviewer 1, we now successfully apply
the same analyses to the fMRI recordings of 100 subjects reading word sequences.

Other comments:

- Rather than using the feature importance delta_R, is it possible to just
    report the parameter estimate E_hat at each time point? What would be the
    downside of this?

It is possible to report $\hat E$ (now relabelled $\hat S$), however they are
interpretable only if the second regression $H$ is not regularized.
We now added a Supplementary Figure to that effect.

Note, however, that E_hat statistics are not in the same unit as the forward H_hat
coefficients (a.k.a betas values in GLM). Consequently, it is not straightforward to
compare these two metrics directly. As a proxy, we thus report their respective
second-level p-values across subjects. The associated figure shows that B2B's $\hat S$
coefficients compare favorably to the Forward H_hat coefficient and the MANOVA
coefficients.

- The authors describe their method as unbiased but then introduce two
    regularization parameters, which will bias the estimated coefficients towards
    zero. While the statement "B2B leads to unbiased (i.e. zeros-centered) scalar
    coefficients for non-causal features" is correct for the mean of those
    coefficients, the variance will also be biased. I think these are important
    criteria to consider.

We agree with this remark. We now clarify the meaning and implication of the
bias in the problem statement:

"Note that $\hat S$ is unbiased, in the sense that it is centered around zero when there is no
effect, only if the second regression $H$ is not regularized. Second-level
statistics testing whether $\hat S$ is superior to 0 are thus only valid if $H$
is not regularized."

- The results of backward regression are not shown in Figure 4, but it would be
    nice to see those results. Is there a possibility to add them to Figure 4 in a
    separate panel?

We now added the decoding performance to a supplementary figure. The latter show
that approximately the same cortical network can be used to decode Word Length
and Word Frequency. It also shows that the Dummy Variable can be decoded from
various brain areas, and from the visual cortex in particular.

- I think the authors could better explain the dimensions of the data (dy
    referring to the number of measurement channels, and dx to the number of
    independent variables). Perhaps the authors could add this nomenclature to their
    graphical illustration? Likewise, for the regression of Y onto X, it is unclear
    if this is a mass-univariate regression or a multivariate regression (i.e.
    taking the error covariance between Ys into account or not). This might be
    obvious to them, but is not necessarily obvious to the reader.

We amended the text in multiple places to clarify this issue.

- I find the choice of the letter E for one of the causal factors unfortunate
    since it is typically reserved for the true error component of a multivariate
    model. Likewise, the choice of N is unfortunate since the letter n has multiple
    meanings in the text: one particular instance of n, or the number of samples.
    This is a very minor comment though.

We now changed $E$ into $S$ for Selection matrix. We also changed the number of
samples into $m$, and keep $N$ as the Noise matrix. Thank you for these suggestions

- Typo line 28: "were" should read "where"?
Yes this is now corrected.

- Line 181: do the authors refer to the Wilcoxon signed-rank test?
Yes, this is now corrected.